---
output: html_notebook
---

<center><h4>Mini research project. <br>
Team: Denys Datsko, Bernaszikovskyi Viacheslav, Hrytsay Maksym</h4></center>

### Intro
Our topic is about chess, because we like to play chess and want to know a little bit more about this game.

We found a good data at https://www.ficsgames.org/download.html
This is a set of many chess games from some internet platdorms

We downloaded the data for the whole 2019 year in blitz mode. There are about 2 million games.

It was in .PGN format, so we made a parser in Python to get all the data in convenient .CSV format.

In the .csv file we have 4 columns representing the rating of a player, whether he made a castling, whether he won and the number of the move, castling was made on.



```{r}
library("BSDA")
install.packages("fitdistrplus")
library("fitdistrplus")
```


### Let us look at the data and plot the information we have for better understanding of it

```{r}
df <- read.csv2("whole_year_2019_blitz.csv", header=TRUE, sep=",")

castling_rate <- data.frame(rating = seq(1000, 2100, by=20), all=0, castling=0)
victories <- data.frame(rating = seq(1000, 2100, by=20), castling=0, wins=0)
overall <- data.frame(rating = seq(1000, 2100, by=20), games=0, wins=0)
```


```{r}
for (i in castling_rate$rating) {
  castling_rate$all[castling_rate$rating == i] = length(df$rating[df$rating %in% seq(i, i + 20)])
  castling_rate$castling[castling_rate$rating == i] = sum(df$castling[df$rating %in% seq(i, i + 20)])
  
  overall$games[overall$rating == i] = castling_rate$all[castling_rate$rating == i]
  
  won <- df[df$rating %in% seq(i, i + 20),]
  won <- as.numeric(won$victory) == 3
  overall$wins[overall$rating == i] = sum(won)
  
  with_castling <- df[as.numeric(df$castling)==1,]
  with_castling <- with_castling[with_castling$rating %in% seq(i, i + 20),]
  
  victories$castling[victories$rating == i] = length(with_castling$rating)
  with_castling <- with_castling[as.numeric(with_castling$victory) == 3,]
  victories$wins[victories$rating == i] <- length(with_castling$rating)
}
```


#### Let us plot the graph of dependency of castling frequency on the player rating


```{r}
plot(x = castling_rate$rating, y = castling_rate$castling / castling_rate$all, type="l")
```

From here we can see, that in general, the bigger rating of the player is, the more popular move castling is.
This can mean, that castling is an efective move, as almost all games by players with high rating are with castling played


```{r}
plot(x = overall$rating, y = overall$wins / overall$games, type="l")
```




```{r}
plot(x = victories$rating, y = victories$wins / victories$castling, type="l")
```






```{r}
func <- victories$wins / victories$castling ~ victories$rating
linear_model = lm(func)
summary(linear_model)

plot(func, col="blue", pch=16)
abline(linear_model, col="red")
```
```{r}
data_wins <- df$victory[as.numeric(df$victory) != 2]
length(df$victory)

data.castling_wins = df$victory[df$castling == 1]
data.castling_wins = data.castling_wins #[as.numeric(data.castling_wins) != 2]


x <- as.numeric(as.character(data_wins))
y = as.numeric(as.character(data.castling_wins))
t.test(x = x, y = y, alternative="two.sided")
```

```{r}
data.no_castling <- df$victory[df$castling==0]
data.no_castling_wins <- data.no_castling #[as.numeric(data.no_castling) != 2]
x <- as.numeric(as.character(data_wins))
y <- as.numeric(as.character(data.no_castling_wins))
t.test(x=x, y=y, alternative="two.sided")
```

```{r}
castling_moves = df$move_number[df$move_number != 0]
hist(castling_moves, breaks=seq(2, 55), probability=T)


descdist(castling_moves)
fit.gamma <- fitdist(castling_moves, "gamma")
plot(fit.gamma)
```

```{r}
shape <- fit.gamma$estimate[1]
rate <- fit.gamma$estimate[2]

ks.test(castling_moves, "pgamma", shape=shape, rate=rate)
```

