---
output:
  html_document:
    df_print: paged
---

<center><h4>Mini research project. <br>
Team: Denys Datsko, Bernaszikovskyi Viacheslav, Hrytsay Maksym</h4></center>

### Intro
Our topic is about chess because we like to play chess and want to know a little bit more about this game.

We found good data at https://www.ficsgames.org/download.html
This is a set of many chess games from some internet platforms

We downloaded the data for the whole 2019 year in blitz mode. There are about 2 million games.

It was in .PGN format, so we made a parser in Python to get all the data in convenient.CSV format.

Parser and the parsed .csv file can be found here: https://github.com/Hryts/castling_research.

In the .csv file we have four columns representing the rating of a player, whether he made a castling, whether he won and the number of the move, castling was made on.


```{r}
library("BSDA")
#install.packages("fitdistrplus")
library("fitdistrplus")
```


### Let us look at the data and plot the information we have for better understanding of it

```{r}
df <- read.csv2("whole_year_2019_blitz.csv", header=TRUE, sep=",")

castling_rate <- data.frame(rating = seq(1000, 2100, by=20), all=0, castling=0)
victories <- data.frame(rating = seq(1000, 2100, by=20), castling=0, wins=0)
overall <- data.frame(rating = seq(1000, 2100, by=20), games=0, wins=0)
```


```{r}
for (i in castling_rate$rating) {
  castling_rate$all[castling_rate$rating == i] = length(df$rating[df$rating %in% seq(i, i + 20)])
  castling_rate$castling[castling_rate$rating == i] = sum(df$castling[df$rating %in% seq(i, i + 20)])
  
  overall$games[overall$rating == i] = castling_rate$all[castling_rate$rating == i]
  
  won <- df[df$rating %in% seq(i, i + 20),]
  won <- as.numeric(won$victory) == 3
  overall$wins[overall$rating == i] = sum(won)
  
  with_castling <- df[as.numeric(df$castling)==1,]
  with_castling <- with_castling[with_castling$rating %in% seq(i, i + 20),]
  
  victories$castling[victories$rating == i] = length(with_castling$rating)
  with_castling <- with_castling[as.numeric(with_castling$victory) == 3,]
  victories$wins[victories$rating == i] <- length(with_castling$rating)
}
```


#### Let us plot the graph of dependency of castling frequency on the player rating


```{r}
plot(x = castling_rate$rating, y = castling_rate$castling / castling_rate$all, type="l", main="Castling frequency", 
     xlab="Rating", ylab="Frequency")
```


From here we can see that in general, the bigger the rating of the player is, the more popular move castling is.
This can mean, that castling is an effective move, as almost all games by players with a high rating are with castling played.


#### Let us plot the graph of dependency of win rate with castling on player`s rating

```{r}
plot(x = victories$rating, y = victories$wins / victories$castling, type="l", 
     main="Win rate with castling", xlab="Rating", ylab="Win rate")
```


It seems this data fits into some linear model. Maybe, the castling is a very powerful move
Let us look at the correlation of win rate with castling and castling frequency


```{r}
cor(victories$wins / victories$castling, castling_rate$castling / castling_rate$all)
```

It seems, that the more player does castling, the more he wins. 
But, let us plot the graph of general win rate (without paying attention to castling) and analyze it.


```{r}
plot(x = overall$rating, y = overall$wins / overall$games, type="l", 
     main="Win rate without castling", xlab="Rating", ylab="Win rate")
```

As we can see here, the graph is almost the same, as the one of win rate with castling.

Let us find their correlation.

```{r}
cor(overall$wins / overall$games, victories$wins / victories$castling)
```

It is almost one, which means that castling almost does not matter for the win rate.

#### It is interesting that win rates with castling have a close to the linear dependency of the rating. Let us test this.

```{r}
func <- victories$wins / victories$castling ~ victories$rating
linear_model = lm(func)
plot(func, col="blue", pch=16, main="Dependency of win rate on rating", xlab="Rating", ylab="Win rate")
abline(linear_model, col="red")
```

The data should fit the LM very well, so let us look at the sumary

```{r}
summary(linear_model)
```
The F-statistic p-value is extremely small, so is B's p-value. Also, Multiple R-squared is very big (0.95 out of 1).
This shows that our data fits into this linear model almost perfectly. So, there is a linear dependency between player's rating and win rate.
#### Why this happens?
In general, at some internet platforms, there are not so many players with a high or low rating, so most of the games
are between players with a big difference in ratings. This leads to the fact that the bigger rating is, the bigger number
of games the player will play with weaker opponents.



### Tests

We can consider all the games, in which one player won (exclude all draws)
This way, all the data can be represented as many Bernoulli random variables (0 if the player lost with probability 1 - p, 1 if he won with probability p)
We have two data sets: all the games and all the games with castling. We know that the unbiased estimator for P is the sample mean, so we can perform a t-test (as the variance is unknown) for estimating, whether two P parameters of variables are equal.

If they are equal, it means, that castling does not influence the game results, if they are not, then castling increases or decreases the win rate (depending on the result)
```{r}
data_wins <- df$victory[as.numeric(df$victory) != 2]
length(df$victory)

data.castling_wins = df$victory[df$castling == 1]
data.castling_wins = data.castling_wins

x <- as.numeric(as.character(data_wins))
y = as.numeric(as.character(data.castling_wins))
t.test(x = x, y = y, alternative="two.sided")
```
The mean of the first data is 0.5, as we analyzed only the data from many games (in which one player wins and the second one loses)
The mean of the second data is 0.509, almost 0.51. This is not so big difference, but for such a big data (almost 4 million elements), this is very significant. This is also approved by the p-value of the test. It is extremely small, so we should reject the null hypothesis. This means that the castling is really a useful move and the one is more likely to win, making the castling.


#### But, what is the average win rate if the player does not do castling?

We will perform the same t-test as in the previous chapter.

```{r}
data.no_castling <- df$victory[df$castling==0]
data.no_castling_wins <- data.no_castling 
x <- as.numeric(as.character(data_wins))
y <- as.numeric(as.character(data.no_castling_wins))
t.test(x=x, y=y, alternative="two.sided")
```
We get 



```{r}
castling_moves = df$move_number[df$move_number != 0]
hist(castling_moves, breaks=seq(2, 55), probability=T)


descdist(castling_moves)
fit.gamma <- fitdist(castling_moves, "gamma")
plot(fit.gamma)
```

```{r}
shape <- fit.gamma$estimate[1]
rate <- fit.gamma$estimate[2]

ks.test(castling_moves, "pgamma", shape=shape, rate=rate)
```
We see that the resulting p-value of the test is extremely small. This means that our data is almost for sure not drawn from the Gamma distribution.

But why is the p-value so small?
That is because for p-value to be larger, the distribution of data of such a size should be extremely close to the expected one with almost no variation.



#### Dependency of success of castling on the number of a move when it was made

We also decided to check if it is more sufficient to make castling earlier or later.
Here is the plot of victory rate after castling in dependence on the move when it was made.

```{r}
move_number_success_rate = data.frame(move_number=seq(1, 30), wins=0, number_of_games=0)

for(i in 1:length(df$move_number)){
  if(as.numeric(df$victory[i]) == 3 && df$move_number[i] <= 30){
    move_number_success_rate$wins[df$move_number[i]] = move_number_success_rate$wins[df$move_number[i]] + 1
  }
  if(df$move_number[i] <= 30){
    move_number_success_rate$number_of_games[df$move_number[i]] = move_number_success_rate$number_of_games[df$move_number[i]] + 1
  }
}

```

```{r}
func2 <- move_number_success_rate$wins/move_number_success_rate$number_of_games ~ move_number_success_rate$move_number
linear_model2 <- lm(func2)
plot(func2, col="blue", pch=16, main="Dependency of win rate on castling move number", xlab="Castling move number", ylab="Win rate")
abline(linear_model2, col="red")
summary(linear_model2)
```
From summary we see, that R-squared value is .7511, which tells us that the linear model is quite relevant.
F-statistic is big enough to judge about dependence.


Here we can see that it is more profitable to castle later. We suppose that it is a loss of tempo to castle early. Besides, the opponent can improve his attack on the side where the king is.
