---
output: html_notebook
---

<center><h4>Mini research project. <br>
Team: Denys Datsko, Bernaszikovskyi Viacheslav, Hrytsay Maksym</h4></center>

### Intro
Our topic is about chess, because we like to play chess and want to know a little bit more about this game.

We found a good data at https://www.ficsgames.org/download.html
This is a set of many chess games from some internet platdorms

We downloaded the data for the whole 2019 year in blitz mode. There are about 2 million games.

It was in .PGN format, so we made a parser in Python to get all the data in convenient .CSV format.

Parser and the parsed .csv file can be found here: https://github.com/Hryts/castling_research

In the .csv file we have 4 columns representing the rating of a player, whether he made a castling, whether he won and the number of the move, castling was made on.



```{r}
library("BSDA")
install.packages("fitdistrplus")
library("fitdistrplus")
```


### Let us look at the data and plot the information we have for better understanding of it

```{r}
df <- read.csv2("whole_year_2019_blitz.csv", header=TRUE, sep=",")

castling_rate <- data.frame(rating = seq(1000, 2100, by=20), all=0, castling=0)
victories <- data.frame(rating = seq(1000, 2100, by=20), castling=0, wins=0)
overall <- data.frame(rating = seq(1000, 2100, by=20), games=0, wins=0)
```


```{r}
for (i in castling_rate$rating) {
  castling_rate$all[castling_rate$rating == i] = length(df$rating[df$rating %in% seq(i, i + 20)])
  castling_rate$castling[castling_rate$rating == i] = sum(df$castling[df$rating %in% seq(i, i + 20)])
  
  overall$games[overall$rating == i] = castling_rate$all[castling_rate$rating == i]
  
  won <- df[df$rating %in% seq(i, i + 20),]
  won <- as.numeric(won$victory) == 3
  overall$wins[overall$rating == i] = sum(won)
  
  with_castling <- df[as.numeric(df$castling)==1,]
  with_castling <- with_castling[with_castling$rating %in% seq(i, i + 20),]
  
  victories$castling[victories$rating == i] = length(with_castling$rating)
  with_castling <- with_castling[as.numeric(with_castling$victory) == 3,]
  victories$wins[victories$rating == i] <- length(with_castling$rating)
}
```


#### Let us plot the graph of dependency of castling frequency on the player rating


```{r}
plot(x = castling_rate$rating, y = castling_rate$castling / castling_rate$all, type="l", main="Castling frequency", 
     xlab="Rating", ylab="Frequency")
```

From here we can see, that in general, the bigger rating of the player is, the more popular move castling is.
This can mean, that castling is an efective move, as almost all games by players with high rating are with castling played


#### Let us plot the graph of dependency of win rate with castling on player`s rating

```{r}
plot(x = victories$rating, y = victories$wins / victories$castling, type="l", 
     main="Win rate with castling", xlab="Rating", ylab="Win rate")
```

It seems, this data fits into some linear model. Maybe, the castling is really a very powerful move
Let us look at the corelation of win rate with castling and castling frequency


```{r}
cor(victories$wins / victories$castling, castling_rate$castling / castling_rate$all)
```
It seems, that the more player does castling, the more he wins. 
But, let us plot the graph of general win rate (without paying attention to castling) and analyze it


```{r}
plot(x = overall$rating, y = overall$wins / overall$games, type="l", 
     main="Win rate without castling", xlab="Rating", ylab="Win rate")
```

As we can see here, the graph is almost the same, as the one of win rate with castling.

Let us find their corelation

```{r}
cor(overall$wins / overall$games, victories$wins / victories$castling)
```

It is almost one, which means, that castling almost does not matter for win rate

#### It is interesting, that win rates with castling have a close to linear dependency of the rating. Let us test this

```{r}
func <- victories$wins / victories$castling ~ victories$rating
linear_model = lm(func)
plot(func, col="blue", pch=16, main="Dependency of win rate on rating", xlab="Rating", ylab="Win rate")
abline(linear_model, col="red")
```

The data should fit the LM very well, so let us look at the sumary

```{r}

summary(linear_model)
```
The F-statistic p-value is extremely small, so is B`s p-value. Also, Multiple R-squared is very big (0.95 out of 1).
This shows, that our data fits into this linear model almost perfectly. So, there is a linear dependency between player`s rating and win rate.

#### Why this happens?
In general, at some internet platforms, there are not so many players with high or low rating, so most of games
are between players with a big difference in ratings. This leads to the fact, that the bigger rating is, the bigger number
of games the player will play with weaker opponents.



### Tests

We can consider all the games, in which one player won (exclude all draws)
This way, all the data can be represented as many Bernoulli random variables (0 if the player lost tieh probability 1 - p , 1 if he won with probability p)
We have two data sets: all the games and all the games with castling. We know, that the unbiased estimator fot P is sample mean, so we can perform a t-test (as variance is unknown) for estimating, whether two P parameters of variables are equal.

If they are equal, it means, that castling does not influence the game results, if they are not, then castling increases or decreases the win rate (depending on the result)

```{r}
data_wins <- df$victory[as.numeric(df$victory) != 2]
length(df$victory)

data.castling_wins = df$victory[df$castling == 1]
data.castling_wins = data.castling_wins

x <- as.numeric(as.character(data_wins))
y = as.numeric(as.character(data.castling_wins))
t.test(x = x, y = y, alternative="two.sided")
```
The mean of the first data is 0.5, as we anlysed only the data from many games (in which one player wins and the second one loses)
The mean of the second data is 0.509, almost 0.51. This is not so bog difference, but for such a big data (almost 4 million elements), this is very significant. This is also approved by the p-value of the test. It is extremely small, so we should reject the null hypotesis. This means, that the castling is really a useful move and the one is more likely to win, making the castling


#### But, what is the average win rate if the player does not do castling?

We will perform the same t-test as in the previous chapter.

```{r}
data.no_castling <- df$victory[df$castling==0]
data.no_castling_wins <- data.no_castling 
x <- as.numeric(as.character(data_wins))
y <- as.numeric(as.character(data.no_castling_wins))
t.test(x=x, y=y, alternative="two.sided")
```
We get 



```{r}
castling_moves = df$move_number[df$move_number != 0]
hist(castling_moves, breaks=seq(2, 55), probability=T)


descdist(castling_moves)
fit.gamma <- fitdist(castling_moves, "gamma")
plot(fit.gamma)
```

```{r}
shape <- fit.gamma$estimate[1]
rate <- fit.gamma$estimate[2]

ks.test(castling_moves, "pgamma", shape=shape, rate=rate)
```

```{r}
move_number_success_rate = data.frame(move_number=seq(1, 30), wins=0, number_of_games=0)

for(i in 1:length(df$move_number)){
  if(as.numeric(df$victory[i]) == 3 && df$move_number[i] <= 30){
    move_number_success_rate$wins[df$move_number[i]] = move_number_success_rate$wins[df$move_number[i]] + 1
  }
  if(df$move_number[i] <= 30){
    move_number_success_rate$number_of_games[df$move_number[i]] = move_number_success_rate$number_of_games[df$move_number[i]] + 1
  }
}
plot(x=move_number_success_rate$move_number, y=move_number_success_rate$wins/move_number_success_rate$number_of_games)

func2 <- move_number_success_rate$wins/move_number_success_rate$number_of_games ~ move_number_success_rate$move_number
linear_model2 <- lm(func2)
plot(func2, col="blue", pch=16, main="Dependency of win rate on castling move number", xlab="Castling move number", ylab="Win rate")
abline(linear_model2, col="red")
summary(linear_model2)
```